A Gaussian process model of human electrocorticographic data

### 2020-02-21 paper/main.tex

Prior work in humans and animals has shown that the spatial profile of the local
field potential differs by frequency band~\citep[e.g., with respect to volume
conductance properties and contribution to the local field
potential;][]{BuzsEtal12, FrieEtal07, CronEtal11}.  For example, lower frequency
components of the local field potential tend to have higher power and extend
further in space than high-frequency components~\citep[e.g., ][]{MillEtal07,
MannEtal09}.  We therefore wondered whether the reconstructions we observed
might be differently weighting or considering the contributions of activity at
different frequency bands.  We therefore examined a range of frequency bands
($\delta$: 2--4 Hz; $\theta$: 4--8 Hz; $\alpha$: 8--12 Hz; $\beta$: 12--30 Hz;
$\gamma_L$: 30--60 Hz; and $\gamma_H$: 60--100 Hz), along with a measure of
broadband power~\citep[defined as the mean height of a linear robust regression
fit in log-log space to the order 4 Morelet wavelet-computed power spectrum at
50 log-spaced frequencies from from 2--100 Hz;][]{MannEtal09}.  We used
second-order Butterworth bandpass filters to compute the activity patterns
within each narrow frequency band.  We then repeated our within-subject and
across-subject cross-validated reconstruction accuracy tests (analogous to
Fig.~\ref{fig:corrmaps}) separately for each frequency band
(Fig.~\ref{fig:frequency}).  (We also carried out a similar analysis on the
Hilbert transform-computed spectral power within each narrow band; see
Fig.~\freqpower.)  Across both datasets, we found that our approach is best at
reconstructing patterns of broadband activity (right-most bars in
Figs.~\ref{fig:frequency}A and D), a correlate of population firing
rate~\citep{MannEtal09}.  We also achieved good reconstruction accuracy within
each narrow frequency band (Figs.~\ref{fig:frequency} and \freqpower).  Activity
at lower frequencies ($\delta$, $\theta$, $\alpha$, and $\beta$) tended to be
reconstructed better than high-frequency patterns ($\gamma_L$ and $\gamma_H$),
with reconstruction accuracy peaking in the $\alpha$ band.  Overall, these
results indicate that our approach is able to accurately recover information
within the 2--100 Hz range, although those reconstructions are most accurate for
low-frequency or broadband patterns.

@@

We also wondered whether the map of information scores might vary as a function
of the spectral components of the activity patterns under consideration.  For
example, low frequency $\delta$ activity in the thalamus has been implicated in
supporting connectivity with cortex~\citep[e.g.,][]{Jone09}. We computed
analogous maps of information scores for each individual frequency band. Across
Datasets 1 and 2 (with the exception of $\alpha$ band activity), we observed
reliable positive correlations between the voxel-by-voxel maps of information
scores ($\delta$: $r = 0.09, p < 10^{-57}$; $\theta$: $r = 0.24, p < 10^{-60}$;
$\alpha$: $r = -0.03, p < 0.001$, $\beta$: $r = 0.02, p - 0.0011$; $\gamma_L$:
$r = 0.1, p < 10^{-67}$; $\gamma_H$: $r = 0.03, p < 10^{-7}$; broadband: $r =
0.21, p < 10^{-297})$.

@@

The frequency-specific information score maps we observed were consistent with
the notion that there is no ``universal'' brain region that  reflects all
activity patterns throughout the rest of the brain.  Rather, each region's
activity patterns appear to be characterized by different spectral profiles, and
the ability to infer full-brain activity patterns at a particular frequency
range depends on the structural and functional connectome specific to that
frequency range (Fig.~\ref{fig:networks}E).  We wondered how the maps we found
might fit in with prior work.  To this end, in addition to examining the
anatomical profiles of each map, we used Neurosynth~\citep{RubiEtal17} to
identify (using meta analyses of the neuroimaging literature) the top five most
common terms associated with each frequency-specific map
(Fig.~\ref{fig:networks}C).  We found that $\delta$ patterns across the brain
were best predicted by regions of ventromedial prefrontal cortex, striatum, and
thalamus (yellow).  These regions are also implicated in modulating $\delta$
oscillations during sleep, and are heavily interconnected with
cortex~\citep[e.g.,][]{AmziSter98}.  The brain areas most predictive of
full-brain $\theta$ patterns were occipital and parietal regions traditionally
linked with visual processing and visual attention (light green). Prior work has
implicated $\theta$ oscillations in these areas in periodic sampling of visual
attention~\citep[e.g.,][]{BuscVanR10}.  We found that full-brain $\alpha$
patterns were best predicted by motor areas (dark green), which also exhibit
$\alpha$ band changes during voluntary movements~\citep[e.g.,][]{JurkEtal06}.
Striatum and thalamus (teal) were most informative about full-brain $\beta$
patterns.  Prior work has implicated striatal $\beta$ activity in sensory and
motor processing~\citep{FeinEtal15} and thalamic $\beta$ activity has been
implicated in modulating widespread $\beta$ patterns across
neocortex~\citep{SherEtal16}. Somatosensory areas (dark blue) were most
informative about full-brain $\gamma_L$ patterns. Prior work has implicated
somatosensory $\gamma_L$ in somatosensory processing and motor
planning~\citep{AyaEtal03}.  Occipital cortex (purple) was most informative
about full-brain $\gamma_H$ patterns.  Occipital $\gamma_H$ has also been linked
with visual processing and reading~\citep{WuEtal11} and the transmission of
visual representations from low-order to higher-order visual
areas~\citep{MatsEtal13}.  Full-brain broadband patterns were best predicted by
inferior parietal cortex precuneus (maroon).  Functional neuroimaging BOLD
responses~\citep{SimoEtal16} and broadband ECoG patterns~\citep{HoneEtal12} in
these default-mode hubs have been implicated in processing context-dependent
representations that unfold over long timescales.

>>>>

Prior work in humans and animals has shown that the spatial profile of the local
field potential differs by frequency band~\citep[e.g., with respect to volume
conductance properties and contribution to the local field
potential;][]{BuzsEtal12, FrieEtal07, CronEtal11}.  For example, lower frequency
components of the local field potential tend to have higher power and extend
further in space than high-frequency components~\citep[e.g., ][]{MillEtal07,
MannEtal09}.  We therefore wondered whether the reconstructions we observed
might be differently weighting or considering the contributions of activity at
different frequency bands.  We therefore examined a range of frequency bands
($\delta$: 2--4 Hz; $\theta$: 4--8 Hz; $\alpha$: 8--12 Hz; $\beta$: 12--30 Hz;
$\gamma_L$: 30--60 Hz; and $\gamma_H$: 60--100 Hz), along with a measure of
broadband power~\citep[defined as the mean height of a linear robust regression
fit in log-log space to the order 4 Morelet wavelet-computed power spectrum at
50 log-spaced frequencies from from 2--100 Hz;][]{MannEtal09}.  We used
second-order Butterworth bandpass filters to compute the activity patterns
within each narrow frequency band.  We then repeated our within-subject and
across-subject cross-validated reconstruction accuracy tests (analogous to
Fig.~\ref{fig:corrmap}) separately for each frequency band
(Fig.~\ref{fig:frequency}).  (We also carried out a similar analysis on the
Hilbert transform-computed spectral power within each narrow band; see
Fig.~\freqpower.)  Across both datasets, we found that our approach is best at
reconstructing patterns of broadband activity (right-most bars in
Figs.~\ref{fig:frequency}A and D), a correlate of population firing
rate~\citep{MannEtal09}.  We also achieved good reconstruction accuracy within
each narrow frequency band (Figs.~\ref{fig:frequency} and \freqpower).  Activity
at lower frequencies ($\delta$, $\theta$, $\alpha$, and $\beta$) tended to be
reconstructed better than high-frequency patterns ($\gamma_L$ and $\gamma_H$),
with reconstruction accuracy peaking in the $\alpha$ band.  Overall, these
results indicate that our approach is able to accurately recover information
within the 2--100 Hz range, although those reconstructions are most accurate for
low-frequency or broadband patterns.

@@

We also wondered whether the map of information scores might vary as a function
of the spectral components of the activity patterns under consideration.  We
computed analogous maps of information scores for each individual frequency
band. Across Datasets 1 and 2 (with the exception of $\alpha$ band activity), we
observed reliable positive correlations between the voxel-by-voxel maps of
information scores ($\delta$: $r = 0.09, p < 10^{-57}$; $\theta$: $r = 0.24, p <
10^{-60}$; $\alpha$: $r = -0.03, p < 0.001$, $\beta$: $r = 0.02, p - 0.0011$;
$\gamma_L$: $r = 0.1, p < 10^{-67}$; $\gamma_H$: $r = 0.03, p < 10^{-7}$;
broadband: $r = 0.21, p < 10^{-297})$.

@@

The frequency-specific information score maps we observed were consistent with
the notion that there is no ``universal'' brain region that  reflects all
activity patterns throughout the rest of the brain.  Rather, each region's
activity patterns appear to be characterized by different spectral profiles, and
the ability to infer full-brain activity patterns at a particular frequency
range depends on the structural and functional connectome specific to that
frequency range (Fig.~\ref{fig:networks}E).  We wondered how the maps we found
might fit in with prior work.  To this end, in addition to examining the
anatomical profiles of each map, we used Neurosynth~\citep{RubiEtal17} to
identify (using meta analyses of the neuroimaging literature) the top five most
common terms associated with each frequency-specific map
(Fig.~\ref{fig:networks}C).  We found that $\delta$ patterns across the brain
were best predicted by regions of ventromedial prefrontal cortex, striatum, and
thalamus (yellow).  These regions are also implicated in modulating $\delta$
oscillations during sleep, and are heavily interconnected with
cortex~\citep[e.g.,][]{AmziSter98}.  The brain areas most predictive of
full-brain $\theta$ patterns were occipital and parietal regions traditionally
linked with visual processing and visual attention (light green). Prior work has
implicated $\theta$ oscillations in these areas in periodic sampling of visual
attention~\citep[e.g.,][]{BuscVanR10}.  We found that full-brain $\alpha$
patterns were best predicted by motor areas (dark green), which also exhibit
$\alpha$ band changes during voluntary movements~\citep[e.g.,][]{JurkEtal06}.
Striatum and thalamus (teal) were most informative about full-brain $\beta$
patterns.  Prior work has implicated striatal $\beta$ activity in sensory and
motor processing~\citep{FeinEtal15} and thalamic $\beta$ activity has been
implicated in modulating widespread $\beta$ patterns across
neocortex~\citep{SherEtal16}. Somatosensory areas (dark blue) were most
informative about full-brain $\gamma_L$ patterns. Prior work has implicated
somatosensory $\gamma_L$ in somatosensory processing and motor
planning~\citep{IharEtal03}.  Occipital cortex (purple) was most informative
about full-brain $\gamma_H$ patterns.  Occipital $\gamma_H$ has also been linked
with visual processing and reading~\citep{WuEtal11} and the transmission of
visual representations from low-order to higher-order visual
areas~\citep{MatsEtal13}.  Full-brain broadband patterns were best predicted by
inferior parietal cortex precuneus (maroon).  Functional neuroimaging BOLD
responses~\citep{SimoEtal16} and broadband ECoG patterns~\citep{HoneEtal12a} in
these default-mode hubs have been implicated in processing context-dependent
representations that unfold over long timescales.
